Introduction to Structural Causal Models

### 2021-06-10 paper_scm.Rmd
stuff

For many research problems, we want to understand the relationship between variables beyond co-occurrence.
The most popular causal model is the structural causal model or in short SCM.
The SCM is an expressive simulator to estimate causal relationships, accounting latent factors. 
The basis of these SCMs is a set of equations, or more precisely assignments, providing functions to derive the conditional probabilities for our model [@hardtrecht].
These assignments describe our variables in our model.
Compared to probabilistic models, where we only specify conditional probabilities, the SCM actually enables the combination of theoretical and observational data.
Conditional probabilities cannot represent latent variables because there is no conditional probability in our observational data for unobserved variables [@pearl2009causality].

History

@@

@@

$\{S\}$ - Smoking
$\{C\}$ - Cancer
$\{E\}$ - Eating

@@

# Graphical Models

-> how to obtain scm

>>>>

For many research problems, we want to understand the relationship between variables beyond co-occurrence.

Research dealing with these question moved from traditional statistics, to ,

The most popular causal model is the structural causal model or in short SCM.
The SCM is an expressive simulator to estimate causal relationships, accounting latent factors. 
The basis of these SCMs is a set of equations, or more precisely assignments, providing functions to derive the conditional probabilities for our model [@hardtrecht].
These assignments describe our variables in our model.
Compared to probabilistic models, where we only specify conditional probabilities, the SCM actually enables the combination of theoretical and observational data.
Conditional probabilities cannot represent latent variables because there is no conditional probability in our observational data for unobserved variables [@pearl2009causality].

@@

The structural causal model consists of a set of equations.
These equations are asymmetric, because they are not bi-directional.
We cannot change the sides of the equation as one can do with regular equations [@pearl2012causal].

@@

$\{S\}$ - Smoking
$\{C\}$ - Cancer
$\{E\}$ - Exercise

Probabilisitic Representation

```{tikz, notwell,  fig.cap ="Probabilistic Model",  fig.align="center", echo =F, fig.show="hold", out.width="50%" }
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
\begin{tikzpicture}[
    sharp corners=2pt,
    inner sep=7pt,
    node distance=3cm,
    >=latex]
\tikzstyle{my node}=[draw, shape = circle, minimum height=1cm,minimum width=1cm]
\node[my node] (S){S};
\node[my node,right =1 cm of S](E){E};
\node[my node] at ($(S)!0.5!(E)-(0pt,1.5cm)$) (C) {C};
\draw[->] (S) -- (C);
\draw[->] (E) -- (C);
\end{tikzpicture}
```

@@

### 2021-06-10 presentation/presentation_video.Rmd
stuff

- Nodes -> Variables (endogenous/exogenous)
- Edges -> relationship (equations)
- Parents/Ancestors/Descendants
- No need to specify exact parametric shape
- Focus on acyclical structures (DAG)
- Estimation back door criterion 
- Test theoretical model structure via causal algorithms to detect structure in data (IC/PC Algo.) 
- More information: [@morgan_winship_2014]

>>>>

- Nodes -> Variables (endogenous/exogenous)
- Edges -> relationship (equations)
- Parents/Ancestors/Descendants
- No need to specify exact parametric shape
- Focus on acyclic structures (DAG)
- Estimation back door criterion 
- Test theoretical model structure via causal algorithms to detect structure in data (IC/PC Algo.) 
- More information: [@morgan_winship_2014]

### 2021-06-10 presentation/scm_short_presentation.Rmd
stuff

- Mechanisms independent/autonomous 
- Set of equations (assignments ':=')
- Observational Data 
- Endogenous and Exogenous Variables
- Latent noise variables 
- Noise terms independent

>>>>

- Mechanisms independent/autonomous 
- Set of equations (assignments $':='$)
- Observational Data 
- Endogenous and Exogenous Variables
- Latent noise variables to simulate joint distribution
- Noise terms $U_1..U_i$ (joint) independence
